### 2020-05-29 writeup/proof_main.tex
add

Given the optimal ratio $\hat v$, let $\hat{M} = \hat{v} \Sigma_1^{1/2}\Sigma_2^{-1/2}$ denote the weighted covariate shift matrix. Denote by ${\hat\lambda}_1\ge {\hat\lambda}_2 \ge \dots \ge {\hat\lambda}_p$ the eigenvalues of $\hat{M}^{\top}\hat{M}$. Let $(\hat a_1, \hat a_2)$ be the solutions to the following deterministic equations
	\be
		 \hat a_1 +  \hat a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad  \hat a_1 + \frac1{\rho_1 + \rho_2}\cdot \frac1p\sum_{i=1}^p \frac{ \hat \lambda_i^2 \hat a_1}{ \hat \lambda_i^2\hat a_1 +  \hat a_2} = \frac{\rho_1}{\rho_1 + \rho_2}.\label{eq_a2} \\
		 \ee
		 After obtaining $(\hat a_1,\hat a_2)$, we can solve the following linear equations to get $(\hat a_3,\hat a_4)$:
\begin{gather}
		\left(\rho_2 \hat a_2^{-2}- \hat b_0\right)\cdot \hat  a_3 - \hat b_1 \cdot \hat a_4
		=\hat b_0, \quad \left(\rho_1\hat a_1^{-2} - \hat b_2  \right)\cdot \hat a_4 - \hat b_1 \cdot \hat a_3 =\hat b_1 .\label{eq_a3} 
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
	\end{gather}
where we denoted
$$\hat b_k:= \frac1{p}\sum_{i=1}^p \frac{\hat \lambda_i^{2k}}{ (\hat a_2 +\hat  \lambda_i^2\hat a_1)^2  },\quad k=0,1,2.$$
Then we introduce the following matrix
$$\Pi = \frac{\rho_1^2}{(\rho_1 + \rho_2)^2}\cdot{\hat M} \frac{(1 + \hat a_3)\id +\hat  a_4 \hat {M}^{\top}\hat {M}}{(\hat a_2 +\hat  a_1 \hat {M}^{\top}\hat {M})^2} \hat {M}^{\top},$$
which is defined in a way such that {\it in certain sense} it is the asymptotic limit of the random matrix 
$$\hat v\Sigma_1^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma_2 (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1^{1/2}.$$ 
We introduce two factors that will appear often in our statements and discussions:
$$\al_-(\rho_1):=(1- \rho_1^{-1/2})^2,\quad \al_+(\rho_1):=(1 + \rho_1^{-1/2})^2.$$ 
In fact, $\al_-^2$ and $\al_+^2$ correspond to the largest and smallest singular values of $Z_1/\sqrt{n_1}$, respectively, as given by the famous Mar{\v c}enko-Pastur law \cite{MP}. In particular, as $\rho_1$ increases, both $\al_-$ and $\al_+$ will converge to 1 and $Z_1/\sqrt{n_1}$ will be more close to an isometry. Finally, we introduce the error term  
$$\delta:=\frac{\al_+^2(\rho_1) - 1 }{\al_-^{2}(\rho_1)} \cdot \kappa^4(\hat M)\cdot \norm{\Sigma_1^{1/2}(\beta_s - \hat{w}\beta_t)}^2,$$
where we recall that $\kappa(\hat M)$ is the condition number of $\hat M$. Note that this factor converges to 0 as $\rho_1\to \infty$.

@@

\begin{theorem}\label{thm_model_shift}
For $i=1,2$, let $Y_i = X_i\beta_i + \varepsilon_i$ be two independent data models, where $X_i=Z_i\Sigma_i^{1/2}\in \R^{n_i\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_i:=n_i/p>1$ being fixed constants, and $\e_i\in \R^{n_i}$ are  random vectors with i.i.d. entries with mean zero, variance $\sigma^2$ and all moments as in \eqref{assmAhigh}.  
	Then with high probability, we have
	\begin{align}
	 	\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL}) \quad \text{ when: } \ \ &\Delta_{\vari} - \Delta_{\beta} \ge  \delta \label{upper}\\
		\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL}) \quad \text{ when: } \ \ &\Delta_{\vari} - \Delta_{\beta} \le -\delta, \label{lower}
	\end{align}
	where
	\begin{align*} %\bigtr{{\Sigma_2^{-1}}}
		\Delta_{\vari} &\define {\sigma^2}\bigbrace{\frac{1}{\rho_2 - 1} -  \frac{1}{\rho_1 + \rho_2}\cdot \frac1p \bigtr{(\hat a_1 \hat M^{\top}\hat M + \hat a_2\id)^{-1}} } \\
		\Delta_{\beta} &\define (\beta_s - \hat{v}\beta_t)^{\top} \Sigma_1^{1/2} \Pi \Sigma_1^{1/2} (\beta_s - \hat{v}\beta_t).
	\end{align*}
\end{theorem}

@@

\begin{lemma}\label{lem_cov_derivative}
	In the setting of Theorem \ref{thm_model_shift}, we have with high probability $1-{\rm o}(1)$,
\begin{equation}\label{lem_cov_derv_eq}
\begin{split}
&\wh w^{2}(n_1+n_2)^2\bignorm{\Sigma_2^{1/2} (\hat{w}X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1 (\beta_s - {w}\beta_t)}^2 \\
&= (\beta_s - {w}\beta_t)^{\top} \Sigma_1^{1/2} {M} \frac{(1 + a_3)\id + a_4 {M}^{\top}{M}}{(a_2 + a_1 {M}^{\top}{M})^2} {M}^{\top} \Sigma_1^{1/2} (\beta_s - \hat{w}\beta_t) +\OO(n^{-1/2+\e}),
\end{split}
\end{equation}
	for any constant $\epsilon>0$. %, where $a_{3,4}$ are found using equations in  \eqref{m35reduced}
\end{lemma}
We will give the proof of this lemma in Section \ref{sec_maintools}.

>>>>

Given the optimal ratio $\hat v$, let $\hat{M} = \hat{v} \Sigma_1^{1/2}\Sigma_2^{-1/2}$ denote the weighted covariate shift matrix. Denote by ${\hat\lambda}_1\ge {\hat\lambda}_2 \ge \dots \ge {\hat\lambda}_p$ the eigenvalues of $\hat{M}^{\top}\hat{M}$. Let $(\hat a_1, \hat a_2)$ be the solutions to the following deterministic equations
	\be
		 \hat a_1 +  \hat a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad  \hat a_1 + \frac1{\rho_1 + \rho_2}\cdot \frac1p\sum_{i=1}^p \frac{ \hat \lambda_i^2 \hat a_1}{ \hat \lambda_i^2\hat a_1 +  \hat a_2} = \frac{\rho_1}{\rho_1 + \rho_2}.\label{eq_a2} \\
		 \ee
		 After obtaining $(\hat a_1,\hat a_2)$, we can solve the following linear equations to get $(\hat a_3,\hat a_4)$:
\begin{gather}
		\left(\rho_2 \hat a_2^{-2}- \hat b_0\right)\cdot \hat  a_3 - \hat b_1 \cdot \hat a_4
		=\hat b_0, \quad \left(\rho_1\hat a_1^{-2} - \hat b_2  \right)\cdot \hat a_4 - \hat b_1 \cdot \hat a_3 =\hat b_1 .\label{eq_a3} 
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
	\end{gather}
where we denoted
$$\hat b_k:= \frac1{p}\sum_{i=1}^p \frac{\hat \lambda_i^{2k}}{ (\hat a_2 +\hat  \lambda_i^2\hat a_1)^2  },\quad k=0,1,2.$$
Then we introduce the following matrix
$$\Pi = \frac{\rho_1^2}{(\rho_1 + \rho_2)^2}\cdot{\hat M} \frac{(1 + \hat a_3)\id +\hat  a_4 \hat {M}^{\top}\hat {M}}{(\hat a_2 +\hat  a_1 \hat {M}^{\top}\hat {M})^2} \hat {M}^{\top},$$
which is defined in a way such that {\it in certain sense} it is the asymptotic limit of the random matrix 
$$\hat v\Sigma_1^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma_2 (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1^{1/2}.$$ 
We introduce two factors that will appear often in our statements and discussions:
$$\al_-(\rho_1):=(1- \rho_1^{-1/2})^2,\quad \al_+(\rho_1):=(1 + \rho_1^{-1/2})^2.$$ 
In fact, $\al_-^2$ and $\al_+^2$ correspond to the largest and smallest singular values of $Z_1/\sqrt{n_1}$, respectively, as given by the famous Mar{\v c}enko-Pastur law \cite{MP}. In particular, as $\rho_1$ increases, both $\al_-$ and $\al_+$ will converge to 1 and $Z_1/\sqrt{n_1}$ will be more close to an isometry. Finally, we introduce the error term  
\be\label{eq_deltaextra}\delta:=\frac{\al_+^2(\rho_1) - 1 }{\al_-^{2}(\rho_1)} \cdot \kappa^4(\hat M)\cdot \norm{\Sigma_1^{1/2}(\beta_s - \hat{w}\beta_t)}^2,\ee
where we recall that $\kappa(\hat M)$ is the condition number of $\hat M$. Note that this factor converges to 0 as $\rho_1\to \infty$.

@@

\begin{theorem}\label{thm_model_shift}
For $i=1,2$, let $Y_i = X_i\beta_i + \varepsilon_i$ be two independent data models, where $X_i=Z_i\Sigma_i^{1/2}\in \R^{n_i\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_i:=n_i/p>1$ being fixed constants, and $\e_i\in \R^{n_i}$ are  random vectors with i.i.d. entries with mean zero, variance $\sigma^2$ and all moments as in \eqref{assmAhigh}.  
	Then with high probability, we have
	\begin{align}
	 	\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL}) \quad \text{ when: } \ \ &\Delta_{\vari} - \Delta_{\beta} \ge  \delta \label{upper}\\
		\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL}) \quad \text{ when: } \ \ &\Delta_{\vari} - \Delta_{\beta} \le -\delta, \label{lower}
	\end{align}
	where
	\begin{align} %\bigtr{{\Sigma_2^{-1}}}
		\Delta_{\vari} &\define {\sigma^2}\bigbrace{\frac{1}{\rho_2 - 1} -  \frac{1}{\rho_1 + \rho_2}\cdot \frac1p \bigtr{(\hat a_1 \hat M^{\top}\hat M + \hat a_2\id)^{-1}} }  \\
		\Delta_{\beta} &\define (\beta_s - \hat{v}\beta_t)^{\top} \Sigma_1^{1/2} \Pi \Sigma_1^{1/2} (\beta_s - \hat{v}\beta_t). 
	\end{align}
\end{theorem}

@@

### 2020-05-29 writeup/transfer.tex
add

\begin{figure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.eps}
		\caption{Model dissimilarity}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.eps}
		\caption{Large source task data size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.eps}
		\caption{Covariate shift}
		\label{fig_covariate}
	\end{subfigure}
	\caption{Three contributing causes of negative transfer:
	(a) As model dissimilarity increases, we observe a transition from positive to negative transfer  (Proposition \ref{prop_dist_transition}).
	(b) As source task data size increases, we observe a transition positive to negative transfer (Proposition \ref{prop_data_size}).
	(c) As covariate shift becomes more severe, test performance gets worse (Proposition \ref{prop_covariate}).}
	\label{fig_model_shift_phasetrans}
\end{figure}

@@

	To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, consider the following family of matrices
	\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \det\bigbrace{ M^\top M} \le \mu^p, \lambda(M) \in [\mu_{\min}, \mu_{\max}]},
	\end{align*}
	where $\mu, \mu_{\min}, \mu_{\max}$ are fixed values that do not grow with $p$.
	Similar to the isotropic model, we assume that $\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_1$ equals $\beta_2$ plus i.i.d. entries with mean $0$ and variance $d^2$.
	The following proposition shows that when $n_1$ is much larger than $n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized at $M = \mu\id$ among all matrices in $\cS_{\mu}$.

\begin{proposition}\label{prop_covariate}
	In the setting described above, assume that $\rho_1 > 3$ and $\rho_2>3$,
and $\|\Sigma_1\|\le C_1$ for some constant $C_1>0$.
	Let $g(M)$ denote the test error of $\hat{\beta}_t^{\MTL}$ when the covariance shift matrix is equal to $M\in\cS_{\mu}$.
	We have that \[ g(\mu\id) \le \bigbrace{1+ \bigo{\frac{\rho_2}{\rho_1} + \frac{1}{\sqrt {\rho_1}}}} \min_{M\in\cS_{\mu}} g(M). \]
\end{proposition}
Proposition \ref{prop_covariate} implies that when $\rho_1\gg \rho_2$, having no covariate shift is the optimal choice for choosing the source task.
%This provides evidence that covariate shift is unfavorable when there are many source task datapoints,

@@

	In Figure \ref{fig_covariate}, we plot the test error of the target task for $n_2 = 4p$ and $n_1$ ranging from $p$ to $20p$.
	Second, we observe the following two phases as we increase $n_1 / p$.
	When $n_1 \le n_2$, having complementary covariance matrices leads to lower test error compared to the case when $\Sigma_1 = \Sigma_2$.
	When $n_1 > n_2$, having complementary covariance matrices leads to higher test error compared to the case when $\Sigma_1 = \Sigma_2$.
	A theoretical justification can be found in Appendix \ref{app_proof_31}.
\end{example}

>>>>

\begin{figure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.eps}
		\caption{Model dissimilarity}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.eps}
		\caption{Large source task data size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.eps}
		\caption{Covariate shift}
		\label{fig_covariate}
	\end{subfigure}
	\caption{Three contributing causes of negative transfer:
	(a) As model dissimilarity increases, we observe a transition from positive to negative transfer  (Proposition \ref{prop_dist_transition}).
	(b) As source task data size increases, we observe a transition positive to negative transfer (Proposition \ref{prop_data_size}).
	(c) As covariate shift becomes more severe (measured by increasing $\lambda$), test performance gets worse (Proposition \ref{prop_covariate}).}
	\label{fig_model_shift_phasetrans}
\end{figure}

@@

	To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, consider the following family of matrices
	\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \det\bigbrace{ M^\top M} \le \mu^p, \lambda(M) \in [\mu_{\min}, \mu_{\max}]},
	\end{align*}
	where $\mu, \mu_{\min}, \mu_{\max}$ are fixed values that do not grow with $p$. We first assume that there is no model bias. 
	Similar to the isotropic model, we assume that $\beta_1=\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$.
	The following proposition shows that when $n_1$ is much larger than $n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized at $M = \mu\id$ among all matrices in $\cS_{\mu}$.

\begin{proposition}\label{prop_covariate}
	In the setting described above, assume that $\rho_1 > 1$ and $\rho_2>1$. 
	%Moreover, up to some rescaling, we can assume that $\|\Sigma_1\|\le C_1$ for some constant $C_1>0$.
	Let $g(M)$ denote the test error of $\hat{\beta}_t^{\MTL}$ when the covariance shift matrix is equal to $M\in\cS_{\mu}$.
	We have that \[ g(\mu\id) \le \bigbrace{1+ \bigo{\frac{\rho_2}{\rho_1}  }} \min_{M\in\cS_{\mu}} g(M). \]
\end{proposition}
Proposition \ref{prop_covariate} implies that when $\rho_1\gg \rho_2$, having no covariate shift is the optimal choice for choosing the source task.
%This provides evidence that covariate shift is unfavorable when there are many source task datapoints,

@@

	In Figure \ref{fig_model_shift_phasetrans} (c), we plot the test error of the target task for $n_2 = 4p$ and $n_1$ ranging from $p$ to $20p$.
	Second, we observe the following two phases as we increase $n_1 / p$.
	When $n_1 \le n_2$, having complementary covariance matrices leads to lower test error compared to the case when $\Sigma_1 = \Sigma_2$.
	When $n_1 > n_2$, having complementary covariance matrices leads to higher test error compared to the case when $\Sigma_1 = \Sigma_2$.
	A theoretical justification can be found in Appendix \ref{app_proof_31}.
\end{example}
